NMRlipids III: Lipid-cholesterol interactions in atomistic resolution molecular dynamics simulations

### 2019-05-08 MANUSCRIPT/manuscript.tex
Update manusript

%\label{}
\section{Introduction}
Details of intermolecular interactions determine the phase
behaviour and details lateral organization of lipid and bilayers
containing cholesterol \cite{ipsen87}.~\todo{Fix Ole's name among the authors of this paper.}
Formation of lateral heterogeneities \cite{kinnunen91}, lipid rafts \cite{simons97}
and superlattices \cite{somerharju09} in cellular membranes have been
suggested to be driven by interactions between cholesterol
and lipids. While detailed experimental information of these interactions
is relatively sparse, atomistic resolution molecular dynamics simulations
have been widely applied to give detailed explanations of lipid bilayer
lateral organization and lipid-cholesterol interactions \cite{rog14,sodt14,??}.
However, the simulations must reproduce the measured experimental details, such as NMR
order parameters and scattering form factors, to be useful in intepretation
of molecular details in lipid bilayer mixtures.

@@

@@

Experimental acyl chain order parameters of pure POPC lipid bilayers
are well reproduced by most force fields (Fig. \ref{OrderParametersCHOL}),
except for the C$_2$ carbon in {\it sn}-1 chain,
which is typically the case in the state of the art lipid force fields (for review see \cite{ollila16}).
%This is also mainly observed in the data
%presented here, except that
However, the order parameters in the beginning of {\it sn}-1 chain are underestimated in the
MACROG simulation and CHARMM36 simulations slightly overestimates the order parameters, also when
used with other than Gromacs simulation packages (section \ref{CHARMMprograms} in the supplementary information).
The absolute values of {\it sn}-1 acyl chain order parameters exhibit approximately linear increase upon addition
of cholesterol also in simulations below equimolar mixture (Figs. \ref{slopesberger}-\ref{slopesslipids310K}).
However, the slope of the increase (Fig. \ref{OrderParametersCHOLchanges}) and order parameters
in equimolar mixtures (Fig. \ref{OrderParametersCHOL}) are overestimated by all force fields.
%For Berger/H\"oltje and MACROG models the overestimation is significant,
%while less severe for CHARMM36 and Slipid models.
The overestimation of order parameters observed in CHARMM without cholesterol and in all force fields
with cholesterol is smaller than the contribution by undulations in large simulations (section \ref{undulations} in the supplementary information).
\begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/slopes.eps}
  \caption{\label{OrderParametersCHOLchanges}
    Slopes of order parameters as a function of cholesterol.
  }
\end{figure}

@@

Visual inspection of form factors (Fig. \ref{FormFactors}) and quantitative quality estimation (Fig. \ref{FormFactorsFITNESS})
suggests that the Berge/Holtje force field gives the best agreement with experiments with large
cholesterol concentration. This is surprising because the same model overestimated acyl chain
order parameter increase upon addition of cholesterol more than CHARMM36 or Slipids simulations
(Figs. \ref{OrderParametersCHOL} and \ref{OrderParametersCHOLchanges}).
The Berger/Holtje model also exhibits the most pronounced decrease in the area per molecule
upon addition of cholesterol (Fig. \ref{alps}).
On the other hand, the electron density profiles from Berger/Holtje simulations are closer
to the result from the SDP model then other MD simulation models for both pure POPC bilayer
and equimolar POPC/cholesterol mixture, especially in the middle of the bilayer
where SDP model gives substantially lower density than MD simulations (Fig. \ref{densities}).
Therefore, the good quality of form factor Berger/Holtje model with large amounts of cholesterol
may rather indicate the importance of this low electron density in the bilayer center,
rather than better quality of lipid-cholesterol interactions in this model.
\begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/FFfitness.eps}
  \caption{\label{FormFactorsFITNESS}
    Fitness of form factors between simulations and experiments.
  }
\end{figure}

@@

\section{Presenting the quality of lipid bilayer MD simulation with binary mixtures}

@@

\begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOLfitness.eps}
  \caption{\label{OrderParametersCHOLfitness}
    Difference between order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
  }
\end{figure*}

>>>>

%\label{}
\section{Introduction}

\todo{We should emphasize the validation aspect of intermolecular interactions in binary lipid mixtures in the introduction.}

Details of intermolecular interactions determine the phase
behaviour and details lateral organization of lipid and bilayers
containing cholesterol \cite{ipsen87}.
Formation of lateral heterogeneities \cite{kinnunen91}, lipid rafts \cite{simons97}
and superlattices \cite{somerharju09} in cellular membranes have been
suggested to be driven by interactions between cholesterol
and lipids. While detailed experimental information of these interactions
is relatively sparse, atomistic resolution molecular dynamics simulations
have been widely applied to give detailed explanations of lipid bilayer
lateral organization and lipid-cholesterol interactions \cite{rog14,sodt14,??}.
However, the simulations must reproduce the measured experimental details, such as NMR
order parameters and scattering form factors, to be useful in intepretation
of molecular details in lipid bilayer mixtures.

@@

\subsection{Quantitative quality measure for lipid bilayer structure}

\todo{To be written once we have agreed on the used measure.}

@@

Experimental acyl chain order parameters of pure POPC lipid bilayers
are well reproduced by most force fields (Fig. \ref{OrderParametersCHOL}),
except for the C$_2$ carbon in {\it sn}-1 chain,
which is typically the case in the state of the art lipid force fields (for review see \cite{ollila16}).
%This is also mainly observed in the data
%presented here, except that
However, the order parameters in the beginning of {\it sn}-1 chain are underestimated in the
MACROG simulation and CHARMM36 simulations slightly overestimates the order parameters, also when
used with other than Gromacs simulation packages (section \ref{CHARMMprograms} in the supplementary information).
The absolute values of {\it sn}-1 acyl chain order parameters exhibit approximately linear increase upon addition
of cholesterol also in simulations below equimolar mixture (Figs. \ref{slopesberger}-\ref{slopesslipids310K}).
However, the slope of the increase (Fig. \ref{OrderParametersCHOLchanges}) and order parameters
in equimolar mixtures (Fig. \ref{OrderParametersCHOL}) are overestimated by all force fields.
%For Berger/H\"oltje and MACROG models the overestimation is significant,
%while less severe for CHARMM36 and Slipid models.
The overestimation of order parameters observed in CHARMM without cholesterol and in all force fields
with cholesterol is smaller than the contribution by undulations in large simulations (section \ref{undulations} in the supplementary information).
\begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/slopes.eps}
  \caption{\label{OrderParametersCHOLchanges}
    Slopes of order parameters as a function of cholesterol.
    Determined by fitting equation $S_{\rm CH}(C_{\rm chol})=k_{\rm OP}C_{\rm chol}+S_{\rm CH}(0)$
    to the data in figures \ref{slopes}-\ref{slopesslipids310K}.
  }
\end{figure}

@@

Visual inspection of form factors (Fig. \ref{FormFactors}) and quantitative quality estimation (Fig. \ref{FormFactorsFITNESS})
suggests that the Berge/Holtje force field gives the best agreement with experiments with large
cholesterol concentration. This is surprising because the same model overestimated acyl chain
order parameter increase upon addition of cholesterol more than CHARMM36 or Slipids simulations
(Figs. \ref{OrderParametersCHOL} and \ref{OrderParametersCHOLchanges}).
The Berger/Holtje model also exhibits the most pronounced decrease in the area per molecule
upon addition of cholesterol (Fig. \ref{alps}).
On the other hand, the electron density profiles from Berger/Holtje simulations are closer
to the result from the SDP model then other MD simulation models for both pure POPC bilayer
and equimolar POPC/cholesterol mixture, especially in the middle of the bilayer
where SDP model gives substantially lower density than MD simulations (Fig. \ref{densities}).
Therefore, the good quality of form factor Berger/Holtje model with large amounts of cholesterol
may rather indicate the importance of this low electron density in the bilayer center,
rather than better quality of lipid-cholesterol interactions in this model.
\begin{figure}[]
  \centering
  \includegraphics[width=8cm]{../FIGS/FFfitness.eps}
  \caption{\label{FormFactorsFITNESS}
    Fitness of form factors between simulations and experiments.
  }
    \todo{Figure to be done properly when the fitness code is finished.
    This is discussed in issues
    https://github.com/NMRLipids/MATCH/issues/65,
    https://github.com/NMRLipids/MATCH/issues/64,
    https://github.com/NMRLipids/MATCH/issues/56, and
    https://github.com/NMRLipids/MATCH/issues/50
  }
\end{figure}

@@

\subsection{Presenting the quality of lipid bilayer MD simulation with binary mixtures}

@@

\begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{../FIGS/OrderParametersCHOLfitness.eps}
  \caption{\label{OrderParametersCHOLfitness}
    Difference between order parameters from simulations and experiments for acyl chains of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
  }
  \todo{Figure to be done properly when the fitness code is finished.
    This is discussed in issues
    https://github.com/NMRLipids/MATCH/issues/65,
    https://github.com/NMRLipids/MATCH/issues/64,
    https://github.com/NMRLipids/MATCH/issues/56, and
    https://github.com/NMRLipids/MATCH/issues/50
  }
\end{figure*}
